\documentclass{memoir}
\begin{document}
%<*"ltType">
resource
%</"ltType">
%<*"ltTopic">
vespersforsunday
%</"ltTopic">
%<*"ltDomain">
en_uk_lash
%</"ltDomain">
%<*"p0001.title.cover">
﻿VESPERS FOR SUNDAYS AND FEASTS
%</"p0001.title.cover">
%<*"p0001.title.doc">
﻿VESPERS
%</"p0001.title.doc">
%<*"p0001.title.toc">
﻿VESPERS
%</"p0001.title.toc">
%<*"p0001.title.header">
﻿VESPERS
%</"p0001.title.header">
%<*"p0001b.title.doc">
﻿FOR SUNDAYS AND FEASTS
%</"p0001b.title.doc">
%<*"p0001b.title.toc">
﻿FOR SUNDAYS AND FEASTS
%</"p0001b.title.toc">
%<*"p0001b.title.header">
﻿FOR SUNDAYS AND FEASTS
%</"p0001b.title.header">
%<*"p0002">
After the Dismissal of the Ninth Hour the Priest and Deacon, having made a metania to the Bishop’s stall, venerated the icons and bowed to the two Choirs, enter the Sanctuary by the side doors. They make three bows to the Holy Table and the Priest kisses the Holy Table and the Gospel.
%</"p0002">
%<*"p0003">
Then, vested in the Epitrachelion, he opens the curtain of the Holy Doors and, standing uncovered in front of the Holy Table, gives the blessing as follows:
%</"p0003">
%<*"p0004">
duplicates.d0199
%</"p0004">
%<*"p0005">
Then the Superior, or the appointed person, reads the Opening Psalm, one of the ‘Gerontika’, as follows:
%</"p0005">
%<*"p0006">
duplicates.d0100
%</"p0006">
%<*"p0007">
duplicates.d0243
%</"p0007">
%<*"p0008">
duplicates.d0240
%</"p0008">
%<*"p0009">
Come, let us worship and fall down before Christ himself, the King and our God.
%</"p0009">
%<*"p0010">
Psalm 103
%</"p0010">
%<*"p0011">
Bless the Lord, my soul! O Lord my God, you have been greatly magnified. You have clothed yourself with thanksgiving and majesty, wrapping yourself in light as in a cloak, stretching out the heavens like a curtain, roofing his upper chambers with waters, placing clouds as his mount, walking on the wings of the wind, making spirits his Angels and a flame of fire his Ministers, establishing the earth on its sure base; it will not be moved to age on age. The deep, like a cloak, is its mantle; waters will stand upon the mountains. At your rebuke they will flee; they will quail at the voice of your thunder. The mountains will rise and the plains descend to the place which you established for them. You fixed a limit that they will not pass, nor will they return to cover the earth. You send out springs into the valleys; waters will run between the mountains. They will give drink to all the beasts of the field; the wild asses will await them to quench their thirst. Beside them the birds of the air will make their dwelling: and sing among the rocks. He waters the mountains from his upper chambers; the earth will be filled from the fruit of your works. He makes grass spring up for the cattle, and green herb for the service of mankind; to bring food out of the earth, and wine makes glad the human heart; to make the face cheerful with oil, and bread strengthens the human heart. The trees of the plain will be satisfied, the cedars of Lebanon that you planted. There the sparrows will build their nests; the heron’s dwelling is at their head. The high mountains are for the deer; rocks a refuge for hares. He made the moon to mark the seasons; the sun knew the hour of its setting. You appointed darkness, and it was night, in which all the beasts of the forest will prowl; young lions roaring to plunder and to seek their food from God. The sun rose and they were gathered together and they will lie down in their dens. Man will go out to his labour; and to his labouring until evening. How your works have been magnified, O Lord. With wisdom you have made them all, and the earth was filled with your creation. See, this great, wide sea; there there are creeping things without number, living creatures small and great. There ships go to and fro; this dragon which you fashioned to sport in it. All things look to you to give them their food in due season. When you give it them, they will gather it.
%</"p0011">
%<*"p0012">
If the Anixantaria are to be sung, the Reader only reads as far as this.
%</"p0012">
%<*"p0013">
When you open your hand all things will be filled with goodness. But when you turn away your face they will be troubled. You will take away their spirit, and they will perish and return to their dust. You will send forth your spirit, and they will be created, and you will renew the face of the earth. May the glory of the Lord endure to the ages. The Lord will rejoice at his works. He looks upon the earth and makes it tremble. He touches the mountains, and they smoke. I will sing to the Lord while I live; I will praise my God while I exist. May my words be pleasing to him. While as for me, I shall rejoice in the Lord. O that sinners might perish from the earth, and the wicked, so that they are no more. Bless the Lord, my soul!
%</"p0013">
%<*"p0014">
And again
%</"p0014">
%<*"p0015">
The sun knew the hour of its setting: you made darkness, and it was night. How your works have been magnified, O Lord. With wisdom you have made them all!
%</"p0015">
%<*"p0016">
duplicates.d0385
%</"p0016">
%<*"p0017">
duplicates.d0206
%</"p0017">
%<*"p0018">
duplicates.d0075
%</"p0018">
%<*"p0019">
While the Opening Psalm is being read the Priest, bareheaded and standing in front of the Holy Table, says quietly the following Prayers at the Lighting of the Lamps, having first blessed the Deacon’s Sticharion and Orarion as usual.
%</"p0019">
%<*"p0020">
duplicates.d0019
%</"p0020">
%<*"p0021">
O Lord, compassionate and merciful, long-suffering and full of mercy, listen to our prayer and attend to the voice of our supplication. Make for us a sign for good. Guide us in your way, to walk in your truth. Make glad our hearts to fear your holy Name, because you are great and do wondrous things. You alone are God, and there is none like you, O Lord, among gods: powerful in mercy and loving in strength to help and to console and to save all who hope in your holy Name.
%</"p0021">
%<*"p0022">
duplicates.d0332
%</"p0022">
%<*"p0023">
duplicates.d0025
%</"p0023">
%<*"p0024">
Lord, do not rebuke us in your anger, nor chastise us in your wrath, but deal with us in accordance with your kindness, physician and healer of our souls. Guide us to the harbour of your will. Enlighten the eyes of our hearts to the knowledge of your truth and grant that the rest of the present day and the whole time of our life may be peaceful and without sin, at the prayers of the holy Mother of God and of all the Saints.
%</"p0024">
%<*"p0025">
For yours is the might and yours is the kingdom and the power and the glory of the Father, the Son and the Holy Spirit, now and for ever, and to the ages of ages. Amen.
%</"p0025">
%<*"p0026">
duplicates.d0029
%</"p0026">
%<*"p0027">
Lord our God, remember us, sinners and your unprofitable servants, as we call upon your holy Name, and do not put us to shame from the expectation of your mercy, but graciously grant us, Lord, all the requests that are for salvation, and count us worthy to love and to fear you from our whole heart, and in all things to do your will.
%</"p0027">
%<*"p0028">
duplicates.d0349
%</"p0028">
%<*"p0029">
duplicates.d0032
%</"p0029">
%<*"p0030">
O Lord, who are praised by the holy Powers with never silent hymns and unceasing songs of glory, fill our mouth with your praise to give majesty to your holy Name, and give us a part and an inheritance with all who fear you in truth and who keep your commandments, at the prayers of the holy Mother of God and of all your Saints.
%</"p0030">
%<*"p0031">
duplicates.d0332
%</"p0031">
%<*"p0032">
duplicates.d0034
%</"p0032">
%<*"p0033">
Lord, Lord, who uphold the universe by your immaculate hand, who are long-suffering towards us all and who repent of evils, remember your acts of compassion and your mercy. Visit us in your loving kindness and grant that for the rest of the present day we may escape the manifold wiles of the evil one, and, by the grace of your All-holy Spirit, keep our life free from assault.
%</"p0033">
%<*"p0034">
By the mercy and love for mankind of your Only-begotten Son, with whom you are blessed, together with your all-holy, good and life-giving Spirit, now and for ever, and to the ages of ages. Amen.
%</"p0034">
%<*"p0035">
duplicates.d0036
%</"p0035">
%<*"p0036">
O God, great and wonderful, who order the universe with inexpressible loving-kindness and rich providence; who have granted us also the blessings of this world and brought us near to the promised Kingdom through the blessings that have been bestowed on us already; who have made us turn aside from every evil during that part of the present day which is now over, grant us also to complete what remains without blame in the presence of your holy glory, as we sing your praise, who alone are our God, good and the Lover of mankind.
%</"p0036">
%<*"p0037">
For you are our God, and to you we give glory, to the Father, the Son and the Holy Spirit, now and for ever, and to the ages of ages. Amen.
%</"p0037">
%<*"p0038">
duplicates.d0038
%</"p0038">
%<*"p0039">
Great and most high God, who alone possess immortality, who dwell in unapproachable light, who fashioned all creation with wisdom, who made the separation between the light and the darkness and who placed the sun to have authority over the day and the moon and the stars to have authority over the night, who have counted us sinners worthy even at this present hour to come into your presence with confession and thanksgiving and to offer you our evening hymn of glory; do you, O Lord who love mankind, direct our prayer like incense before you and accept it as a savour of sweet fragrance. Grant us that the present evening and the coming night may be peaceful, clothe us with weapons of light, deliver us from every night-time fear and from every deed that walks in darkness. And give us sleep, which you have bestowed on us for our rest in our weakness, freed from every fantasy of the devil. Yes, Master of all things, giver of blessings, may we also be filled with compunction on our beds and call to mind your Name in the night, and enlightened by meditation on your commandments may we rise with gladness of soul to give glory to your loving-kindness, as we bring to your compassion supplications and entreaties on behalf of our own sins and those of all your people. At the prayers of the holy Mother of God visit them with mercy.
%</"p0039">
%<*"p0040">
duplicates.d0349
%</"p0040">
%<*"p0041">
When the Opening Psalm is completed the Priest, in front of the Holy Table, or the Deacon in his usual place outside the Holy Doors, says the:
%</"p0041">
%<*"p0042">
duplicates.d0559
%</"p0042">
%<*"p0043">
duplicates.d0251
%</"p0043">
%<*"p0044">
duplicates.d0504
%</"p0044">
%<*"p0045">
duplicates.d0773
%</"p0045">
%<*"p0046">
duplicates.d0585
%</"p0046">
%<*"p0047">
duplicates.d0251
%</"p0047">
%<*"p0048">
duplicates.d0318
%</"p0048">
%<*"p0049">
duplicates.d0319
%</"p0049">
%<*"p0050">
duplicates.d0326
%</"p0050">
%<*"p0051">
duplicates.d0291
%</"p0051">
%<*"p0052">
duplicates.d0302
%</"p0052">
%<*"p0053a">
For our Country 
%</"p0053a">
%<*"p0053b">
, our President 
%</"p0053b">
%<*"p0053c">
, Our Defense Forces, and all in civil authority, let us pray to the Lord.
%</"p0053c">
%<*"p0054">
duplicates.d0325
%</"p0054">
%<*"p0055">
duplicates.d0293
%</"p0055">
%<*"p0056">
duplicates.d0327
%</"p0056">
%<*"p0057">
duplicates.d0304
%</"p0057">
%<*"p0058">
duplicates.d0463
%</"p0058">
%<*"p0059">
duplicates.d0246
%</"p0059">
%<*"p0060">
duplicates.d0773
%</"p0060">
%<*"p0061">
duplicates.d1054
%</"p0061">
%<*"p0062">
duplicates.d0779
%</"p0062">
%<*"p0063">
duplicates.d0331
%</"p0063">
%<*"p0064">
duplicates.d0773
%</"p0064">
%<*"p0065">
duplicates.d0100
%</"p0065">
%<*"p0066">
After the Litany, except on the Feasts of the Lord, the appointed Kathisma of the Psalter is read by one Reader.
%</"p0066">
%<*"p0067">
The order of reading the psalms outside of Lent
%</"p0067">
%<*"p0068">
Saturday vespers: Kathisma 1 (psalms 1-8)
%</"p0068">
%<*"p0069">
Sunday matins: Kathisma 2(psalms 9-17), 3(psalms 18-24)
%</"p0069">
%<*"p0070">
Sunday vespers: none
%</"p0070">
%<*"p0071">
Monday matins: Kathisma 4(psalms 25-32), 5(psalms 33-37)
%</"p0071">
%<*"p0072">
Monday vespers: Kathisma 6 (psalms 38-46)
%</"p0072">
%<*"p0073">
Tuesday matins: Kathisma 7(psalms 47-55), 8(psalms 56-64)
%</"p0073">
%<*"p0074">
Tuesday vespers: Kathisma 9 (psalms 65-70)
%</"p0074">
%<*"p0075">
Wednesday matins: Kathisma 10(psalms 71-77), 11(psalms 78-85)
%</"p0075">
%<*"p0076">
Wednesday vespers: Kathisma 12 (psalms 86-91)
%</"p0076">
%<*"p0077">
Thursday matins: Kathisma 13(psalms 92-101), 14(psalms 102-105)
%</"p0077">
%<*"p0078">
Thursday vespers: Kathisma 15 (psalms 106-109)
%</"p0078">
%<*"p0079">
Friday matins: Kathisma 19(psalms 135-143), 20(psalms 144-150)
%</"p0079">
%<*"p0080">
Friday vespers: Kathisma 18 (psalms 120-134)
%</"p0080">
%<*"p0081">
Saturday matins: Kathisma 16(psalms 110-118), 17(psalm 119)
%</"p0081">
%<*"p0082">
Short Litany
%</"p0082">
%<*"p0083">
duplicates.d0251
%</"p0083">
%<*"p0084">
duplicates.d0058
%</"p0084">
%<*"p0085">
duplicates.d0773
%</"p0085">
%<*"p0086">
duplicates.d0575
%</"p0086">
%<*"p0087">
duplicates.d0251
%</"p0087">
%<*"p0088">
duplicates.d0463
%</"p0088">
%<*"p0089">
duplicates.d0773
%</"p0089">
%<*"p0090">
duplicates.d0575
%</"p0090">
%<*"p0091">
duplicates.d0251
%</"p0091">
%<*"p0092">
duplicates.d0246
%</"p0092">
%<*"p0093">
duplicates.d0773
%</"p0093">
%<*"p0094">
duplicates.d1054
%</"p0094">
%<*"p0095">
duplicates.d0779
%</"p0095">
%<*"p0096">
For yours is the might, and yours the kingdom, the power and the glory, of the Father, the Son and the Holy Spirit, now and for ever, and to the ages of ages.
%</"p0096">
%<*"p0097">
duplicates.d0773
%</"p0097">
%<*"p0098">
duplicates.d0100
%</"p0098">
%<*"p0099a">
Then we begin
%</"p0099a">
%<*"p0099b">
Lord, I have cried
%</"p0099b">
%<*"p0099c">
in the appropriate tone, while the Deacon censes the Sanctuary and the whole Church. If the Priest has to cense he does so vested in the Phelonion and Epitrachelion. He blesses the incense with the words,
%</"p0099c">
%<*"p0100">
duplicates.d0199
%</"p0100">
%<*"p0101">
He censes the holy Table from the four sides and the whole sanctuary. Then coming out by the north door he censes the holy icons (x3), the people and the whole church in the customary manner and returns to the sanctuary by the south door.
%</"p0101">
%<*"p0102">
Psalm 140
%</"p0102">
%<*"p0103">
 Lord, I have cried to you, hear me; hear me, O Lord. Lord, I have cried to you, hear me. Give heed to the voice of my supplication when I cry to you. Hear me, O Lord.
%</"p0103">
%<*"p0104">
 Let my prayer be directed like incense before you; the lifting up of my hands be an evening sacrifice. Hear me, O Lord.
%</"p0104">
%<*"p0105">
 Set a guard, O Lord, on my mouth, and a strong door about my lips.
%</"p0105">
%<*"p0106">
 Do not incline my heart to evil words; to make excuses for my sins.
%</"p0106">
%<*"p0107">
 With those who work iniquity, let me not unite with their elect.
%</"p0107">
%<*"p0108">
 The just will chastise me with mercy and reprove me; but let not the oil of sinners anoint my head.
%</"p0108">
%<*"p0109">
 For yet my prayer shall be in their pleasures; their judges have been swallowed up near the rock.
%</"p0109">
%<*"p0110">
[B] They will hear my words for they are sweet. As a clod of earth is crushed upon the ground, their bones have been scattered at the mouth of Hell.
%</"p0110">
%<*"p0111">
[A] For my eyes look to you, O Lord, my Lord. I have hoped in you, do not take away my soul.
%</"p0111">
%<*"p0112">
 Keep me from the snare that they have hidden for me, and from the traps of evil-doers.
%</"p0112">
%<*"p0113">
 The sinners will fall into their own net. I am alone until I pass by.
%</"p0113">
%<*"p0114">
Psalm 141
%</"p0114">
%<*"p0115">
 With my voice I cried to the Lord. With my voice I entreated the Lord.
%</"p0115">
%<*"p0116">
 I will pour out my entreaty before him, and tell him all my trouble.
%</"p0116">
%<*"p0117">
 When my spirit was faint, you knew my path.
%</"p0117">
%<*"p0118">
 In the way where I walked they had hidden a snare for me.
%</"p0118">
%<*"p0119">
 I looked to my right hand and saw, but there was none who knew me.
%</"p0119">
%<*"p0120">
 Escape is gone from me, and there is none who seeks for my soul.
%</"p0120">
%<*"p0121">
 I cried to you, Lord, I said, ‘You are my hope; my portion in the land of the living.’
%</"p0121">
%<*"p0122">
 Give heed to my supplication, for I am brought very low.
%</"p0122">
%<*"p0123">
 Deliver me from those who persecute me, for they are too strong for me.
%</"p0123">
%<*"p0124a">
For 10 Verses, or Stichera 
%</"p0124a">
%<*"p0124b">
[
%</"p0124b">
%<*"p0124c">
On Saturday evening
%</"p0124c">
%<*"p0124d">
]
%</"p0124d">
%<*"p0124e">
.
%</"p0124e">
%<*"p0125a">
10.
%</"p0125a">
%<*"p0125b">
Bring my soul out of prison, that I may confess your name.
%</"p0125b">
%<*"p0126a">
9.
%</"p0126a">
%<*"p0126b">
The just will await me, until you reward me.
%</"p0126b">
%<*"p0127a">
For 8 Verses 
%</"p0127a">
%<*"p0127b">
[
%</"p0127b">
%<*"p0127c">
Always at a full Vigil
%</"p0127c">
%<*"p0127d">
]
%</"p0127d">
%<*"p0127e">
.
%</"p0127e">
%<*"p0128">
Psalm 129
%</"p0128">
%<*"p0129a">
8.
%</"p0129a">
%<*"p0129b">
Out of the depths I have cried to you, O Lord. Lord hear my voice.
%</"p0129b">
%<*"p0130a">
7.
%</"p0130a">
%<*"p0130b">
Let your ears be attentive, to the voice of my supplication.
%</"p0130b">
%<*"p0131">
For 6 Verses.
%</"p0131">
%<*"p0132a">
6.
%</"p0132a">
%<*"p0132b">
If you, Lord, should mark iniquities, Lord, who will stand? But there is forgiveness with you.
%</"p0132b">
%<*"p0133a">
5. 
%</"p0133a">
%<*"p0133b">
For your name’s sake I have waited for you, O Lord. My soul has waited on your word. My soul has hoped in the Lord.
%</"p0133b">
%<*"p0134">
duplicates.d0286
%</"p0134">
%<*"p0135a">
4.
%</"p0135a">
%<*"p0135b">
From the morning watch until night, from the morning watch, let Israel hope in the Lord.
%</"p0135b">
%<*"p0136a">
3.
%</"p0136a">
%<*"p0136b">
For with the Lord there is mercy, and with him plentiful redemption, and he will redeem Israel from all his iniquities.
%</"p0136b">
%<*"p0137">
Psalm 116
%</"p0137">
%<*"p0138a">
2.
%</"p0138a">
%<*"p0138b">
Praise the Lord, all you nations. Praise him all you peoples.
%</"p0138b">
%<*"p0139a">
1.
%</"p0139a">
%<*"p0139b">
For his mercy has been mighty towards us, and the truth of the Lord endures to the ages.
%</"p0139b">
%<*"p0140">
Glory… Both now…
%</"p0140">
%<*"p0141">
During the Doxastikon or Theotokion, which on Saturday evening is always the first of the Tone, the Priest, vested in the Phelonion, and the Deacon make the Entrance with the censer. The Priest blesses the incense as above. They come out through the north door, the Deacon leading, and stop opposite the Holy Doors. The Deacon says,
%</"p0141">
%<*"p0142">
duplicates.d0551
%</"p0142">
%<*"p0143">
The Priest bows his head and says, in a low voice, the
%</"p0143">
%<*"p0144">
Prayer of the Entrance
%</"p0144">
%<*"p0145">
At evening, at morning and at midday we praise, bless and give thanks, and we pray to you, Master of all things, Lord who love mankind: Direct our prayer before you like incense, and do not incline our hearts to words or thoughts of evil, but deliver us from all that hunt down our souls. For our eyes look to you, O Lord, our Lord, and we have hoped in you.
%</"p0145">
%<*"p0146">
duplicates.d0332
%</"p0146">
%<*"p0147">
Then the Deacon says,
%</"p0147">
%<*"p0148">
Master, bless the holy entrance.
%</"p0148">
%<*"p0149">
And the Priest blesses the Entrance with the words:
%</"p0149">
%<*"p0150">
Blessed is the entrance of your holy ones, always, now and for ever, and to the ages of ages.
%</"p0150">
%<*"p0151">
duplicates.d0251
%</"p0151">
%<*"p0152">
duplicates.d0100
%</"p0152">
%<*"p0153">
The Deacon censes the holy icons and the People and waits for the end of the troparion. When it is finished, he raises the censer, making the sign of the Cross, and exclaims:
%</"p0153">
%<*"p0154">
duplicates.d1145
%</"p0154">
%<*"p0155">
And they enter the sanctuary through the Holy Doors, the one with the censer censing the holy Table.
%</"p0155">
%<*"p0156">
And we sing
%</"p0156">
%<*"p0157">
Thanksgiving at the Lighting of the Lamps
%</"p0157">
%<*"p0158">
An ancient poem, or, as some say, by the martyr Athenogenes.
%</"p0158">
%<*"p0159">
O joyful Light of the holy glory of the immortal, heavenly, holy, blessed Father, O Jesus Christ. Now that we have come to the setting of the sun and see the evening light, we sing the praise of God, Father, Son and Holy Spirit. It is right at all times to hymn you with holy voices, Son of God, giver of life. Therefore the world glorifies you.
%</"p0159">
%<*"p0160">
In Athonite use this hymn is always read as one of the Gerontika, except when a number of Priests take part in the Entrance.
%</"p0160">
%<*"p0161a">
Deacon (facing the People)
%</"p0161a">
%<*"p0161b">
Prokeimenon of the evening.
%</"p0161b">
%<*"p0162">
duplicates.d0697
%</"p0162">
%<*"p0163">
duplicates.d1085
%</"p0163">
%<*"p0164">
duplicates.d0903
%</"p0164">
%<*"p0165a">
Verse 1: 
%</"p0165a">
%<*"p0165b">
The Lord has clothed and girded himself with power
%</"p0165b">
%<*"p0166">
duplicates.d0903
%</"p0166">
%<*"p0167a">
Verse 2:
%</"p0167a">
%<*"p0167b">
He has established the world, which will not be shaken.
%</"p0167b">
%<*"p0168">
The Lord is King: * he has clothed himself with glory.
%</"p0168">
%<*"p0169">
On Sunday evening.
%</"p0169">
%<*"p0170">
duplicates.d1090
%</"p0170">
%<*"p0171">
Come, bless the Lord: all you servants of the Lord. 
%</"p0171">
%<*"p0172a">
Verse:
%</"p0172a">
%<*"p0172b">
You that stand in the house of the Lord, in the courts of the house of our God.
%</"p0172b">
%<*"p0173">
Come, bless the Lord: * all you servants of the Lord.
%</"p0173">
%<*"p0174">
On Monday evening.
%</"p0174">
%<*"p0175">
duplicates.d1080
%</"p0175">
%<*"p0176">
When I call to the Lord: * he will hear me.
%</"p0176">
%<*"p0177a">
Verse: 
%</"p0177a">
%<*"p0177b">
Answer me when I call, O God of my justice.
%</"p0177b">
%<*"p0178">
On Tuesday evening.
%</"p0178">
%<*"p0179">
duplicates.d1068
%</"p0179">
%<*"p0180">
Your mercy, Lord, will follow me: * all the days of my life.
%</"p0180">
%<*"p0181a">
Verse:
%</"p0181a">
%<*"p0181b">
The Lord is my shepherd, therefore can I lack nothing.
%</"p0181b">
%<*"p0182">
On Wednesday evening.
%</"p0182">
%<*"p0183">
duplicates.d1084
%</"p0183">
%<*"p0184">
Save me, O God, by the power of your name: * and vindicate me by your might.
%</"p0184">
%<*"p0185a">
Verse:
%</"p0185a">
%<*"p0185b">
Hear my prayer, O God, and listen to the words of my mouth.
%</"p0185b">
%<*"p0186">
On Thursday evening
%</"p0186">
%<*"p0187">
duplicates.d1085
%</"p0187">
%<*"p0188">
My help comes from the Lord: * who made both heaven and earth.
%</"p0188">
%<*"p0189a">
Verse:
%</"p0189a">
%<*"p0189b">
I have lifted up my eyes to the hills, from which my help will come.
%</"p0189b">
%<*"p0190">
On Friday evening.
%</"p0190">
%<*"p0191">
duplicates.d1088
%</"p0191">
%<*"p0192">
O God you are my helper: * your mercy will go before me.
%</"p0192">
%<*"p0193a">
Verse:
%</"p0193a">
%<*"p0193b">
Deliver me from my enemies, O God, ransom me from those that rise up against me.
%</"p0193b">
%<*"p0194">
[On certain great feasts there follow Readings from the Old or New Testament.]
%</"p0194">
%<*"p0195">
duplicates.d0558
%</"p0195">
%<*"p0196">
duplicates.d0251
%</"p0196">
%<*"p0197">
duplicates.d0535
%</"p0197">
%<*"p0198">
duplicates.d0773
%</"p0198">
%<*"p0199">
duplicates.d0575
%</"p0199">
%<*"p0200">
duplicates.d0251
%</"p0200">
%<*"p0201">
duplicates.d0564
%</"p0201">
%<*"p0202">
duplicates.d0439
%</"p0202">
%<*"p0203">
duplicates.d0773
%</"p0203">
%<*"p0204a">
Lord, have mercy.
%</"p0204a">
%<*"p0204b">
(Three times. And so after the remaining petitions)
%</"p0204b">
%<*"p0205">
duplicates.d0251
%</"p0205">
%<*"p0206">
duplicates.d0085
%</"p0206">
%<*"p0207a">
Also we pray for our Country 
%</"p0207a">
%<*"p0207b">
, the President 
%</"p0207b">
%<*"p0207c">
, our Defence Forces and all in authority.
%</"p0207c">
%<*"p0208a">
Also we pray for mercy, life, peace, health, salvation, visitation, pardon and forgiveness of sins for the servants of God, all devout and Orthodox Christians, those who dwell in or visit this city and parish, the wardens and members of this church and their families; 
%</"p0208a">
%<*"p0208b">
[
%</"p0208b">
%<*"p0208c">
and for the servants of God 
%</"p0208c">
%<*"p0208d">
and
%</"p0208d">
%<*"p0208e">
(N) (Here the Deacon may name those for whom he has been asked to pray)
%</"p0208e">
%<*"p0208f">
, and all who have asked for our prayers, unworthy though we are.
%</"p0208f">
%<*"p0208g">
]
%</"p0208g">
%<*"p0209a">
Also we pray for the blessed and ever-remembered founders of this holy church, and for all our departed brothers and sisters, Orthodox believers, who have gone to their rest before us and who here and in all the world lie asleep in the Lord; 
%</"p0209a">
%<*"p0209b">
[
%</"p0209b">
%<*"p0209c">
and for the servants of God 
%</"p0209c">
%<*"p0209d">
and
%</"p0209d">
%<*"p0209e">
(N) (Here the Deacon may name those for whom he has been asked to pray)
%</"p0209e">
%<*"p0209f">
, and that they may be pardoned all their offences, both voluntary and involuntary.
%</"p0209f">
%<*"p0209g">
]
%</"p0209g">
%<*"p0210">
duplicates.d0094
%</"p0210">
%<*"p0211">
duplicates.d0779
%</"p0211">
%<*"p0212">
For you, O God, are merciful and love mankind, and to you we give glory, to the Father and to the Son and to the holy Spirit, now and for ever, and to the ages of ages.
%</"p0212">
%<*"p0213">
duplicates.d0773
%</"p0213">
%<*"p0214">
duplicates.d0100
%</"p0214">
%<*"p0215">
After the Litany say this prayer, one of the Gerontika.
%</"p0215">
%<*"p0216">
Grant, Lord, to keep us this evening without sin. Blessed are you, Lord, the God of our fathers, and praised and glorified is your name to the ages. Amen.
%</"p0216">
%<*"p0217">
Let your mercy, Lord be upon us, as we have hoped in you.
%</"p0217">
%<*"p0218">
Blessed are you, Lord: teach me your statutes.
%</"p0218">
%<*"p0219">
Blessed are you, Master: make me understand your statutes.
%</"p0219">
%<*"p0220">
Blessed are you, Holy One: enlighten me with your statutes.
%</"p0220">
%<*"p0221">
Lord, your mercy is for ever; do not scorn the work of your hands. To you praise is due, to you song is due, to you glory is due, to the Father, and to the Son, and to the Holy Spirit, now and for ever, and to the ages of ages. Amen.
%</"p0221">
%<*"p0222">
duplicates.d0558
%</"p0222">
%<*"p0223">
duplicates.d0251
%</"p0223">
%<*"p0224">
Let us complete our evening prayer to the Lord.
%</"p0224">
%<*"p0225">
duplicates.d0773
%</"p0225">
%<*"p0226a">
Lord, have mercy.
%</"p0226a">
%<*"p0226b">
(And so after each of the following petitions)
%</"p0226b">
%<*"p0227">
duplicates.d0251
%</"p0227">
%<*"p0228">
duplicates.d0462
%</"p0228">
%<*"p0229">
That the whole evening may be perfect, holy, peaceful and sinless, let us ask of the Lord.
%</"p0229">
%<*"p0230">
duplicates.d0773
%</"p0230">
%<*"p0231a">
Grant this, O Lord.
%</"p0231a">
%<*"p0231b">
(And so after each of the following petitions)
%</"p0231b">
%<*"p0232">
duplicates.d0251
%</"p0232">
%<*"p0233">
duplicates.d0107
%</"p0233">
%<*"p0234">
duplicates.d0760
%</"p0234">
%<*"p0235">
duplicates.d1044
%</"p0235">
%<*"p0236">
duplicates.d0858
%</"p0236">
%<*"p0237">
duplicates.d0041
%</"p0237">
%<*"p0238">
duplicates.d0246
%</"p0238">
%<*"p0239">
duplicates.d0773
%</"p0239">
%<*"p0240">
duplicates.d1054
%</"p0240">
%<*"p0241">
duplicates.d0779
%</"p0241">
%<*"p0242">
duplicates.d0348
%</"p0242">
%<*"p0243">
duplicates.d0773
%</"p0243">
%<*"p0244">
duplicates.d0100
%</"p0244">
%<*"p0245a">
Priest (blessing the People)
%</"p0245a">
%<*"p0245b">
Peace to all.
%</"p0245b">
%<*"p0246">
duplicates.d0773
%</"p0246">
%<*"p0247">
duplicates.d0142
%</"p0247">
%<*"p0248">
duplicates.d0251
%</"p0248">
%<*"p0249">
duplicates.d0541
%</"p0249">
%<*"p0250">
duplicates.d0773
%</"p0250">
%<*"p0251">
duplicates.d1054
%</"p0251">
%<*"p0252">
The Priest, in a low voice, says the:
%</"p0252">
%<*"p0253">
duplicates.d0777
%</"p0253">
%<*"p0254">
Lord our God, who bowed the heavens and came down for the salvation of the human race, look upon your servants and upon your inheritance, for to you, the fearful Judge who love mankind, your servants have bowed their heads and inclined their necks, not waiting for any human help, but awaiting your mercy and looking for your salvation. Guard them at every moment, during both the present evening and the approaching night, from every foe, from every hostile operation of the devil and from vain thoughts and evil desires.
%</"p0254">
%<*"p0255a">
(Aloud)
%</"p0255a">
%<*"p0255b">
Blessed and glorified be the might of your kingdom, of the Father and of the Son and of the holy Spirit, now and for ever, and to the ages of ages.
%</"p0255b">
%<*"p0256">
duplicates.d0773
%</"p0256">
%<*"p0257">
duplicates.d0100
%</"p0257">
%<*"p0258">
After the Litany the Aposticha, between which we say the verses of the Feast, if there are any.
%</"p0258">
%<*"p0259">
duplicates.d0697
%</"p0259">
%<*"p0260a">
Verse 1.
%</"p0260a">
%<*"p0260b">
The Lord is King, he has clothed himself with glory; the Lord has clothed and girded himself with power.
%</"p0260b">
%<*"p0261a">
Verse 2.
%</"p0261a">
%<*"p0261b">
He has established the world, which will not be shaken.
%</"p0261b">
%<*"p0262a">
Verse 3.
%</"p0262a">
%<*"p0262b">
Holiness befits your house, O Lord, to length of days.
%</"p0262b">
%<*"p0263">
On the completion of the Aposticha, we say the
%</"p0263">
%<*"p0264">
Song of Symeon who Received God (Luke 2,29),
%</"p0264">
%<*"p0265">
one of the Gerontika
%</"p0265">
%<*"p0266">
duplicates.d0779
%</"p0266">
%<*"p0267">
Now, Master, you let your servant depart in peace, according to your word; for my eyes have seen your Salvation, which you have prepared before the face of all peoples, a Light to bring revelation to the nations, and the Glory of your people Israel.
%</"p0267">
%<*"p0268">
duplicates.d0801
%</"p0268">
%<*"p0269">
duplicates.d0470
%</"p0269">
%<*"p0270">
duplicates.d0398
%</"p0270">
%<*"p0271">
All-holy Trinity, have mercy on us. Lord, cleanse us from our sins. Master, pardon our iniquities. Holy One, visit and heal our infirmities for your name's sake.
%</"p0271">
%<*"p0272">
duplicates.d0586
%</"p0272">
%<*"p0273">
duplicates.d0397
%</"p0273">
%<*"p0274">
Our Father, in heaven, may your name be hallowed, your kingdom come, your will be done on earth as in heaven. Give us today our daily bread, and forgive us our debts, as we forgive our debtors. And lead us not into temptation, but deliver us from the evil one.
%</"p0274">
%<*"p0275">
duplicates.d0779
%</"p0275">
%<*"p0276">
duplicates.d0358
%</"p0276">
%<*"p0277">
duplicates.d0801
%</"p0277">
%<*"p0278">
duplicates.d0100
%</"p0278">
%<*"p0279a">
The Apolytikion of the day, 
%</"p0279a">
%<*"p0279b">
Glory. Both now
%</"p0279b">
%<*"p0279c">
, and the Theotokion in the same Tone.
%</"p0279c">
%<*"p0280">
[If there is an Artoklasia it takes place here.]
%</"p0280">
%<*"p0281">
duplicates.d0779
%</"p0281">
%<*"p0282">
duplicates.d1143
%</"p0282">
%<*"p0283">
duplicates.d0801
%</"p0283">
%<*"p0284">
duplicates.d0184
%</"p0284">
%<*"p0285a">
Priest (from inside the Sanctuary)
%</"p0285a">
%<*"p0285b">
Blessed is He Who Is, Christ our true God, always, now and for ever, and to the ages of ages.
%</"p0285b">
%<*"p0286">
duplicates.d0801
%</"p0286">
%<*"p0287">
duplicates.d0104
%</"p0287">
%<*"p0288">
duplicates.d0779
%</"p0288">
%<*"p0289">
duplicates.d0631
%</"p0289">
%<*"p0290">
duplicates.d0801
%</"p0290">
%<*"p0291">
duplicates.d0431
%</"p0291">
%<*"p0292">
duplicates.d0952
%</"p0292">
%<*"p0293">
duplicates.d0405
%</"p0293">
%<*"p0294">
duplicates.d0801
%</"p0294">
%<*"p0295a">
Glory. Both now. Lord, have mercy. 
%</"p0295a">
%<*"p0295b">
Give the blessing.
%</"p0295b">
%<*"p0296">
And the Priest, lowering the Phelonion ,continues:
%</"p0296">
%<*"p0297a">
May [
%</"p0297a">
%<*"p0297b">
on Saturday evening
%</"p0297b">
%<*"p0297c">
he who rose from the dead,] Christ our true God, at the prayers of his most pure and holy Mother; of the holy, glorious and all-praised Apostles; of Saint 
%</"p0297c">
%<*"p0297d">
[
%</"p0297d">
%<*"p0297e">
the patron of the church
%</"p0297e">
%<*"p0297f">
], of Saint
%</"p0297f">
%<*"p0297g">
, whose memory we celebrate; of the holy and righteous forebears of God, Joachim and Anne and of all the Saints, have mercy on us and save us, for he is good and loves mankind.
%</"p0297g">
%<*"p0298">
Turning to the East he says,
%</"p0298">
%<*"p0299">
duplicates.d1051
%</"p0299">
%<*"p0300">
duplicates.d0773
%</"p0300">
%<*"p0301">
duplicates.d0100
%</"p0301">
%<*"p0302">
duplicates.d0047
%</"p0302">
%<*"p0303">
On Sunday evening in Lent we sing the following Troparia instead of the usual Apolytikion, making a full prostration at each of them, except the last.
%</"p0303">
%<*"p0304">
duplicates.d1084
%</"p0304">
%<*"p0305a">
Virgin Mother of God, hail Mary full of grace, the Lord is with you. Blessed are you among women, and blessed is the fruit of your womb, for you have given birth to the Saviour of our souls.
%</"p0305a">
%<*"p0305b">
(Prostration)
%</"p0305b">
%<*"p0306a">
Baptist of Christ, remember us all, that we may be delivered from our transgressions: for you have been given grace to intercede for us.
%</"p0306a">
%<*"p0306b">
(Prostration)
%</"p0306b">
%<*"p0307">
duplicates.d0415
%</"p0307">
%<*"p0308a">
Pray for us, holy Apostles and all you Saints, that we may be delivered from dangers and afflictions: for in you we have gained fervent advocates with the Saviour.
%</"p0308a">
%<*"p0308b">
(Prostration)
%</"p0308b">
%<*"p0309">
duplicates.d0210
%</"p0309">
%<*"p0310">
We take refuge beneath your compassion, Mother of God: do not despise our petitions in trouble: but rescue us from dangers, only pure, only blessed one.
%</"p0310">
%<*"p0311">
duplicates.d0801
%</"p0311">
%<*"p0312">
duplicates.d0583
%</"p0312">
%<*"p0313">
duplicates.d0397
%</"p0313">
%<*"p0314">
Greater in honour than the Cherubim, and beyond compare more glorious than the Seraphim, without corruption you gave birth to God the Word. Truly the Mother of God we magnify you.
%</"p0314">
%<*"p0315">
Lord, have mercy. Lord, have mercy. Lord, have mercy. Holy father, give the blessing.
%</"p0315">
%<*"p0316">
duplicates.d0779
%</"p0316">
%<*"p0317">
duplicates.d0191
%</"p0317">
%<*"p0318">
duplicates.d0801
%</"p0318">
%<*"p0319">
duplicates.d0100
%</"p0319">
%<*"p0320">
And we say for the Rulers:
%</"p0320">
%<*"p0321">
Heavenly King, establish our rulers: strengthen the faith: calm the nations: make the world peaceful: guard well this holy church: assign our brothers and sisters who have gone before us to the tents of the righteous, and accept us in repentance and confession, as you are good and love mankind.
%</"p0321">
%<*"p0322">
Then three full Prostrations with the Prayer of Saint Ephrem the Syrian once only and then the Dismissal, as above.
%</"p0322">
%<*"p0323">
The Prayer of Saint Ephrem The Syrian
%</"p0323">
%<*"p0324a">
Lord and Master of my life, do not give me a spirit of sloth, idle curiosity, love of power and useless chatter.
%</"p0324a">
%<*"p0324b">
(Prostration)
%</"p0324b">
%<*"p0325a">
Rather accord to me, your servant, a spirit of sobriety, humility, patience and love. 
%</"p0325a">
%<*"p0325b">
(Prostration)
%</"p0325b">
%<*"p0326a">
Yes, Lord and King, grant me to see my own faults and not to condemn my brother; for you are blessed to the ages of ages. Amen.
%</"p0326a">
%<*"p0326b">
(Prostration)
%</"p0326b">
%<*"p0327.title.toc">
NINTH HOUR
%</"p0327.title.toc">
%<*"p0327.title.doc">
NINTH HOUR
%</"p0327.title.doc">
%<*"p0327.title.header">
NINTH HOUR
%</"p0327.title.header">
%<*"p0328">
Vested in the Epitrachelion, the Priest, standing on the Solea, the Holy Doors and the curtain being closed, begins:
%</"p0328">
%<*"p0329">
duplicates.d0199
%</"p0329">
%<*"p0330a">
Reader [in Athonite use the Superior]
%</"p0330a">
%<*"p0330b">
Amen.
%</"p0330b">
%<*"p0331">
duplicates.d0412
%</"p0331">
%<*"p0332">
Heavenly King, Paraclete, Spirit of truth, present everywhere, filling all things, Treasury of blessings and Giver of life, come and dwell in us, cleanse us from every stain, and, O Good One, save our souls.
%</"p0332">
%<*"p0333">
duplicates.d0801
%</"p0333">
%<*"p0334">
duplicates.d0466
%</"p0334">
%<*"p0335">
duplicates.d0398
%</"p0335">
%<*"p0336">
duplicates.d0071
%</"p0336">
%<*"p0337">
duplicates.d0397
%</"p0337">
%<*"p0338">
duplicates.d0736
%</"p0338">
%<*"p0339">
duplicates.d0779
%</"p0339">
%<*"p0340">
duplicates.d0358
%</"p0340">
%<*"p0341">
duplicates.d0801
%</"p0341">
%<*"p0342a">
Amen. Lord, have mercy. 
%</"p0342a">
%<*"p0342b">
Glory. Both now.
%</"p0342b">
%<*"p0343">
duplicates.d0243
%</"p0343">
%<*"p0344">
duplicates.d0240
%</"p0344">
%<*"p0345">
duplicates.d0239
%</"p0345">
%<*"p0346">
3 Metanias, then the Psalms.
%</"p0346">
%<*"p0347">
Psalm 83
%</"p0347">
%<*"p0348">
How beloved are your dwellings, O Lord of powers; my soul longs and faints for the courts of the Lord. My heart and my flesh rejoiced in the living God. For the sparrow has found itself a house, and the dove a nest for herself, where she may place her nestlings: your altars, O Lord of powers, my King and my God. Blessed are those who dwell in your house, they will praise you to ages of ages. Blessed the one whose help is from you; who has placed ascents in his heart, to the valley of weeping, to the place he has set. For the lawgiver will give blessings; they will go from strength to strength, the God of gods will appear in Sion. Lord God of powers hear my prayer, give ear, O God of Jacob. O God our shield see; and look on the face of your Christ. For one day in your courts is far better than thousands. I chose to be cast aside in the house of my God, rather than to dwell in the tents of sinners. Lord God of powers, blessed is the one who hopes in you.
%</"p0348">
%<*"p0349">
Psalm 84
%</"p0349">
%<*"p0350">
You were well pleased with your land, O Lord, you turned away the captivity of Jacob. You forgave your people their iniquities, you covered all their sins. You made all your anger cease, you turned away from the anger of your rage. Turn us back again, O God of our salvation, and turn away your rage from us. Would you be angry with us for ever? Or will you prolong your wrath from generation to generation? O God, when you have turned us back you will make us live, and your people will rejoice in you. Lord, show us your mercy, and grant us your salvation. I will hear what the Lord will say in me; for he will speak peace to his people, to his holy ones, to those who turn their hearts back to him. But his salvation is near those who fear him, that glory may dwell in our land. Mercy and truth have met, righteousness and peace have kissed. Truth has dawned from the earth, and righteousness has leaned down from heaven. For the Lord will give goodness, and our earth will give its fruit. Righteousness will walk before him, and set his steps in the way.
%</"p0350">
%<*"p0351">
Psalm 85
%</"p0351">
%<*"p0352">
Incline your ear, O Lord, and hear me, for I am poor and needy. Guard my soul, for I am holy; O my God, save your servant who hopes in you. Have mercy on me, Lord, for I shall cry to you all the day; give joy to the soul of your servant, for I have lifted up my soul to you. For you, O Lord, are good and gentle, and full of mercy to all who call on you. Give ear, O Lord, to my prayer; and attend to the voice of my supplication. In the day of my trouble I called to you, for you have heard me. There is none like you among the gods, O Lord; none whose works are as yours. All the nations whom you made will come and bow down before you, O Lord, and glorify your name. For you are great and do wondrous things; you alone are God. Guide me, Lord, in your way, and I shall walk in your truth; make my heart glad to fear your name. I shall give you thanks, O Lord, my God, with my whole heart; and I shall glorify your name for ever. For your mercy is great towards me, and you have delivered my soul from the lowest Hell. O God, the lawless rose up against me, and an assembly of mighty ones sought my soul, and did not set you before them. And you, Lord, are pitying and merciful, long-suffering and full of mercy and true. Look on me and have mercy on me; give your might to your servant and save the son of your maidservant. Make with me a sign for good, and let those who hate me see it and be shamed; for you, Lord, have helped me and comforted me.
%</"p0352">
%<*"p0353">
duplicates.d0111
%</"p0353">
%<*"p0354">
Make with me a sign for good, and let those who hate me see it and be shamed; for you, Lord, have helped me and comforted me.
%</"p0354">
%<*"p0355">
duplicates.d0386
%</"p0355">
%<*"p0356">
duplicates.d0206
%</"p0356">
%<*"p0357a">
Alleluia, Alleluia, Alleluia. Glory to you, O God. 
%</"p0357a">
%<*"p0357b">
Lord, have mercy. 
%</"p0357b">
%<*"p0358">
duplicates.d0386
%</"p0358">
%<*"p0359">
Then the Apolytikion of the day. If there are two feasts, the first Apolytikion is said before Glory and the second after it.
%</"p0359">
%<*"p0360">
duplicates.d0211
%</"p0360">
%<*"p0361">
O Good One, born of a Virgin for our sakes and who endured crucifixion, who despoiled death by death and as God revealed resurrection, do not despise those whom you fashioned with your own hand; show your love for mankind, O Merciful; accept the Mother of God who bore you, as she intercedes for us, and save, O Saviour, a people in despair.
%</"p0361">
%<*"p0362">
For your holy name’s sake do not finally reject us, do not annul your covenant, do not take your mercy from us for the sake of Abraham, your beloved, and for the sake of Isaac, your servant, and Israel, your holy one.
%</"p0362">
%<*"p0363">
duplicates.d0470
%</"p0363">
%<*"p0364">
duplicates.d0398
%</"p0364">
%<*"p0365">
duplicates.d0071
%</"p0365">
%<*"p0366">
duplicates.d0397
%</"p0366">
%<*"p0367">
duplicates.d0736
%</"p0367">
%<*"p0368">
duplicates.d0779
%</"p0368">
%<*"p0369">
duplicates.d0358
%</"p0369">
%<*"p0370">
duplicates.d0801
%</"p0370">
%<*"p0371">
duplicates.d0100
%</"p0371">
%<*"p0372">
Then the Kontakion of the day. On Sunday the Ypakoï.
%</"p0372">
%<*"p0373">
duplicates.d0583
%</"p0373">
%<*"p0374">
At every time and at every hour, in heaven and on earth worshipped and glorified, Christ God, long-suffering, great in mercy, great in compassion, loving the just and merciful to sinners, calling all to salvation by the promise of the good things to come; do you, Lord, yourself accept our entreaties at this hour, and direct our lives to your commandments. Sanctify our souls, purify our bodies, correct our thoughts, cleanse our ideas and deliver us from every distress, evil, and pain. Wall us about with your holy Angels, that protected and guided by their host we may reach the unity of the faith and the knowledge of your unapproachable glory; for you are blessed to the ages of ages. Amen.
%</"p0374">
%<*"p0375a">
Lord, have mercy.
%</"p0375a">
%<*"p0375b">
Glory. Both now. Greater in honour than the Cherubim, and beyond compare more glorious than the Seraphim, without corruption you gave birth to God the Word, truly the Mother of God we magnify you. In the name of the Lord, bless, Father.
%</"p0375b">
%<*"p0376">
duplicates.d0779
%</"p0376">
%<*"p0377">
May God take pity on us and bless us, and show the light of his countenance on us and have mercy on us.
%</"p0377">
%<*"p0378">
Then the Superior says this
%</"p0378">
%<*"p0379">
Prayer of Saint Basil the Great
%</"p0379">
%<*"p0380">
Master, Lord Jesus Christ, our God, who have long endured our transgressions, and brought us to this hour in which, hanging on the life-giving tree, you showed the good Thief the way into Paradise and destroyed death by death, have mercy also on us sinners and your unworthy servants. For we have sinned and transgressed, and are not worthy to raise our eyes and look on the height of heaven, because we have abandoned the way of your justice and walked in the will of our hearts. But we implore your unbounded goodness: spare us, O Lord, according to the multitude of your mercy, and save us for your holy name’s sake, for our days have been wasted in vanity. Rescue us from the hand of our opponent, forgive us our sins, slay our carnal will, so that we, having put off the old man, may put on the new, and live for you, our Master and Benefactor; and that thus following your precepts we may reach eternal rest, where those who rejoice have their dwelling. For you are the true joy and gladness of those who love you, Christ our God, and to you we give glory, together with your Father who has no beginning, and your all-holy, good and life-giving Spirit, now and for ever, and to the ages of ages. Amen.
%</"p0380">
%<*"p0381">
duplicates.d0779
%</"p0381">
%<*"p0382">
Glory to you, Christ God, our hope. Glory to you.
%</"p0382">
%<*"p0383">
duplicates.d0801
%</"p0383">
%<*"p0384">
Glory to the Father, and to the Son, and to the Holy Spirit; both now and for ever, and to the ages of ages. Amen. Lord, have mercy. Lord, have mercy. Lord, have mercy. Give the blessing.
%</"p0384">
%<*"p0385">
duplicates.d0779
%</"p0385">
%<*"p0386a">
May Christ our true God, at the prayers of his most pure and holy Mother; the intercessions of the holy, glorious and all-praised Apostles; of Saint
%</"p0386a">
%<*"p0386b">
N. [the patron of the church]
%</"p0386b">
%<*"p0386c">
; of Saint whose memory we keep today, and of all the Saints, have mercy on us and save us, for he is good and loves mankind.
%</"p0386c">
%<*"p0387">
Through the prayers of our holy fathers, Lord Jesus Christ have mercy on us.
%</"p0387">
%<*"p0388">
duplicates.d0801
%</"p0388">
%<*"p0389">
duplicates.d0100
%</"p0389">
